---
title: "Project 1"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(writexl)
library(dplyr)
library(class)
library(stringr)
library(readxl)
library(rworldmap)
library(glmnet)
library(MASS)
library(caret)
library(plyr)
library(e1071)
# Heads up when we switch trade code we need to set our path directories so do it up here 
# so it's easier to see and so we don't have to keep chanigng our file path in the code below
setwd("/Users/alexvidal/Desktop/Spring2024Classes/CMDA_4654/CMDA_Project_1/")
```


```{r, include=FALSE}
file_path <- "cleaned_data.xlsx"
cleaned_data <- read_excel(file_path)
cleaned_data["InternetTrafficPercofResidentsWithInternetAccess.y"] <- NULL
cleaned_data["InternetTraffic12.31.2020.y"] <- NULL
cleaned_data["InternetUsersNumberOfInternetUsers.y"] <- NULL
cleaned_data["Year"] <- NULL


# Changing name for some of the longer column names
names(cleaned_data)[names(cleaned_data) == "InternetTrafficPercofResidentsWithInternetAccess"]<-"Percent People Connected"
names(cleaned_data)[names(cleaned_data) == "InternetTraffic12.31.2020"] <- "Internet Traffic"
names(cleaned_data)[names(cleaned_data) == "InternetUsersNumberOfInternetUsers"] <- "Internet Users Amount"

data1 <- cleaned_data
names(cleaned_data) <- str_replace_all(names(cleaned_data), c(" " = "_"))

# Change data types in economic columns from char io doubles
cleaned_data <- cleaned_data %>% mutate_at(c(3:15), as.numeric)
worldMap <- joinCountryData2Map(cleaned_data, joinCode="NAME", nameJoinColumn="Country")
```

Data Overview {data-orientation=rows}
==================

### Our Data
```{r}
knitr::kable(head(data1))
```



Rows
----------------


### Data Introduction

Our data comprises of two data sets. One data set is from the Index of Economic freedom, (https://www.heritage.org/index/pages/all-country-scores) site that scores the economic freedom of different countries based on particular metrics such as property rights, tax burden, labor freedom ,etc and will calculate an overall score to show how 'free' a countries economic opportunities and such are. The scores are from 0-100 where 100 means the country is the most 'free' in that area of analysis while 0 means no freedom for that particular category. The next data set that we have is the internet usage by country data set found here (https://worldpopulationreview.com/country-rankings/internet-users-by-country). It consists of 3 main variables the total internet users per country, internet traffic in Terabytes and the percentage of internet users in the country. We have combined both of these data sets into one larger data set to do our analysis which we will talk about in the abstract section. 

Here is a link to our two data sets https://drive.google.com/drive/folders/12HrlXJR_UdA65qzBq8JsycShSF3KdxXM?usp=sharing





### Data Dictionary 

Country - The country name

Region - The region the country is in

Overall_Score - The overall Economic Index Score

Property_Rights - The score for the property Rights a country has

Government_Integrity - Score based on how non-corrupt government is

Judicial_Effectiveness - Score for how just judicial system is

Tax_Burden - Score that measures income and corporate tax as a percentage of GDP      

Government_Spending - Score that measures government spending

Fiscal_Health - Score of debt burden on country

Business_Freedom - Scores the environment for business to succeed

Labor_Freedom - Scores the various aspects of the legal and regulatory framework of a country’s labor market

Monetary_Freedom - Scores inflation with an assessment of various government activities that distort prices

Trade_Freedom - Score that measures  the extent of tariff and nontariff barriers that affect imports and exports of goods and services

Investment_Freedom - Score that measures how constrained investing is

Financial_Freedom - Score that measures banking efficiency and a measure of independence from government control and interference in the financial sector

Internet_Users_Amount - The amount of internet users in a country, phone, laptop etc.

Internet_Traffic - The amount of Terabytes of internet traffic a country has

Percent_People_Connected - The percent of people in the country connected to internet. Can alse be more than 100% because of large heavy tourism for that country artificially increasing users. 

Abstract
=====================

### Abstract 

We have a few goals we intend to reach in analyzing this data set. First one being we are interested to find any relationship between particular Economic Freedom parameters that could help predict not only the percentage of internet users for a country as well as what that relationship looks like. We also intend on finding a way to classify a countries region based on particular Economic Freedom Indexes and their respective internet user percentage and see what relationships or differences different regions of the world would have with those parameters. We will use a multitude of classification algorithms from kNN to Naive Bayes and use two multivariate regression models and see the pros and cons of each respective model in our results tab. 

Below we have two plots showing Percentage of Internet users and Ecnonomic Freedom Score and we can already see some patterns between these two variables. We will of course need ot do more analysis to fully flesh out the relationships the two data parameters may have.



Rows
--------------------


```{r, plot1, fig.width=15, fig.height=6}
par(mfrow=c(1,2))
mapCountryData(worldMap, nameColumnToPlot ="Percent_People_Connected", mapTitle="Percent of People Connected to Internet")
mapCountryData(worldMap, nameColumnToPlot ="Overall_Score", mapTitle="Overall Economic Freedom Score")
```


Multivariate Regression {data-orientation=columns}
=====================


```{r, include=FALSE}
parameters <- c("", "Government_Spending", "Business_Freedom", "Monetary_Freedom", "Investment_Freedom", "")
fullModel <- lm(cbind(Percent_People_Connected, Overall_Score) ~Government_Spending+Business_Freedom+Monetary_Freedom+Investment_Freedom, data = cleaned_data)

values <- fullModel$coefficients
newdata <- data.frame(Tax_Burden =50, Government_Spending=50, Business_Freedom=50, Monetary_Freedom=50, Investment_Freedom=50, Financial_Freedom=50)


```

### Parameter Values
```{r}
knitr::kable(fullModel$coefficients, col.names = c("Percent People Connected","Overall Score"))
```



### Description
For Multivariate regression we decided we wanted to see if we can predict the percent of people that are connected to the internet along with the country's overall economic freedom score based on the country's Government spending, business freedom score, monetary freedom score and investment freedom score. The reasoning behind this was just the interest in finding if there was any relation between all of these parameters.

We also plotted the the specific relationship between our predictors and response variable. We noticed that business freedom correlated highest with percent of people connected to the internet. We also found (unsurprisingly) many of the economic freedom index scores were correlated with the overall economic score except for government spending. We are not sure why that is. 



Rows {data-orientation=rows}
---------------------

```{r plot2, fig.width=3, fig.height=2}
ggplot(data=cleaned_data, aes(y=Percent_People_Connected))+
  geom_point(aes(x=Government_Spending), color='blue')+
  labs(x="Score", y="% People using Internet", title = "Government Spending")
ggplot(data=cleaned_data, aes(y=Percent_People_Connected))+
  geom_point(aes(x=Business_Freedom), , color='green')+
  labs(x="Score", y="% People using Internet", title = "Business Freedom")
ggplot(data=cleaned_data, aes(y=Percent_People_Connected))+
  geom_point(aes(x=Monetary_Freedom), color='yellow')+
  labs(x="Score", y="% People using Internet", title = "Monetary Freedom")
ggplot(data=cleaned_data, aes(y=Percent_People_Connected))+
  geom_point(aes(x=Investment_Freedom), color='red')+
  labs(x="Score", y="% People using Internet", title = "Investment Freedom")
```

Rows {data-orientation=rows}
---------------------

```{r plot3, fig.width=3, fig.height=2}
ggplot(data=cleaned_data, aes(y=Overall_Score))+
  geom_point(aes(x=Government_Spending), color='blue')+
  labs(x="Score", y="Overall Score", title = "Government Spending")
ggplot(data=cleaned_data, aes(y=Overall_Score))+
  geom_point(aes(x=Business_Freedom), , color='green')+
  labs(x="Score", y="Overall Score", title = "Business Freedom")
ggplot(data=cleaned_data, aes(y=Overall_Score))+
  geom_point(aes(x=Monetary_Freedom), color='yellow')+
  labs(x="Score", y="Overall Score", title = "Monetary Freedom")
ggplot(data=cleaned_data, aes(y=Overall_Score))+
  geom_point(aes(x=Investment_Freedom), color='red')+
  labs(x="Score", y="Overall Score", title = "Investment Freedom")
```






Ridge Regression {data-orientation=rows}
=====================

### Description
We'll use Percent People Connected as the response variable and the Financial Freedom, Government Spending, Government Integrity as predictors. To perform ridge regression, we use glmnet function to fit the ridge regression with alpha = 0 and find the best lambda. Then we can calculate the R squared pf the model on the training data, the R squared of the model is 43.68%, the best model was able to explain only 43.68% of the variation in the response values of the training data.The goodness of fit of the model is relatively low, these three predictors are not good enough to predict the Percent People Connect of a country.



### Results


Rows {data-orientation=columns}
---------------------

```{r}
# Set up the predictors matrix, the chose predictors variables are Financial Freedom, Government Spending and Government Intergrity
predictors <- c("Financial_Freedom", "Government_Spending", "Government_Integrity")
X <- as.matrix(cleaned_data[, predictors]) 

#set the variable we want to analyze the Percent People Connected
y <- cleaned_data$Percent_People_Connected

# Cross validation and plot
ridge_model <- glmnet(x = X, y = y, alpha = 0)
cv_ridge <- cv.glmnet(x = X, y = y, alpha = 0)
plot(cv_ridge)
```


```{r}
# Fit ridge regression model with min lambda
best_lambda <- cv_ridge$lambda.min
#best_lambda #print out the best lambda
ridge <- glmnet(x = X, y = y, alpha = 0, lambda = best_lambda)

#produce Ridge trace plot
plot(ridge_model, xvar = "lambda")
```


```{r}
# Make prediction based on the ridge regression model
prediction <- predict(ridge_model, s = best_lambda, newx = X)

# Plot the prediction
plot(y,prediction, xlab = "Actual Percent", ylab = "Predicted Percent", main = "Actual Percent vs. Predicted Percent")
model <- lm(prediction~y)
lines(y, predict(model))
#find SST and SSE
sst <- sum((y - mean(y))^2)
sse <- sum((prediction - y)^2)

#find R-Squared
rsq <- 1 - sse/sst
#rsq

```

LOESS Fit {data-orientation=rows}
=====================
### Description
We are using the LOESS to fit a curve to the relationship between variables in pair, Financial Freedom and Percent People Connected, Government Spending and Percent People Connected, Tax Burden and Percent People Connected, Property Rights and Percent People Connected. We can see that only Financial Freedom and Property Rights have a clearly positive relationship with Percent People Connected. This is consistent with the expectation that people of a country have more financial freedom and property rights have more possibility to access to the Internet.




### Results

{data-orientation=rows}
--------------------

```{r plot4, fig.width=7, fig.height=4}
# Fit LOESS curve for financial freedom vs. percent people connected
loess_fit <- loess(Percent_People_Connected ~ cleaned_data$Financial_Freedom, data = cleaned_data)
predicted <- predict(loess_fit)
# Plot data and LOESS fit
ggplot(cleaned_data, aes(x = Financial_Freedom, y = Percent_People_Connected)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(x = "Financial Freedom", y = "Percent People Connected", title = "Financial Freedom vs. Percent People Connected")

# Fit LOESS curve for government spending vs. percent people connected
loess_fit <- loess(Percent_People_Connected ~ cleaned_data$Government_Spending, data = cleaned_data)
predicted <- predict(loess_fit)
# Plot data and LOESS fit
ggplot(cleaned_data, aes(x = Government_Spending, y = Percent_People_Connected)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(x = "Government Spending", y = "Percent People Connected", title = "Government Spending vs. Percent People Connected")
```


```{r plot5, fig.width=7, fig.height=4}
# Fit LOESS curve for tax burden vs. percent people connected
loess_fit <- loess(Percent_People_Connected ~ cleaned_data$Tax_Burden, data = cleaned_data)
predicted <- predict(loess_fit)
# Plot data and LOESS fit
ggplot(cleaned_data, aes(x = Tax_Burden, y = Percent_People_Connected)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(x = "Tax Burden", y = "Percent People Connected", title = "Tax Burden vs. Percent People Connected")

# Fit LOESS curve for tax burden vs. percent people connected
loess_fit <- loess(Percent_People_Connected ~ cleaned_data$Property_Rights, data = cleaned_data)
predicted <- predict(loess_fit)
# Plot data and LOESS fit
ggplot(cleaned_data, aes(x = Property_Rights, y = Percent_People_Connected)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(x = "Property Rights", y = "Percent People Connected", title = "Property Rights vs. Percent People Connected")

```




Description {data-navmenu="KNN Classification" data-orientation=columns}
=====================


### Description
For K-Nearest Neighbors Classification we decided to try and figure out based on Overall Economic Freedom Score and percentage of internet users how you can classify what region that country belongs to. We already have our data classified based on region and some of the regions are larger than others making this knn classification more biased to those countries. We wish to see if the possible relationships that arise in trying to classify the region, like would certain regions would cluster in certain areas like say, higher Monetary Freedom Score index and high internet connection percentage. 

In our testing we decided the other parameter besides percent of people connected to the internet to classify regions with would be the Monetary Freedom Score. For whatever reason this gives our highest accuracy score of 0.5510204 without adding any more variables to be classified with. One possible explanation could be that when compared with percentage of people connected to the internet, the Monetary Freedom Score has the least looking correlation as seen in our multivariate regression tab. 

The plot shows the section that the training set created when it was being trained on with 70% of the whole data set as its training set. Each colored section represents the region's region for classification. So for example Sub-Saharran Africa has a large region that is mainly clustered with a low percentage of people connected to the internet but a fairly high range of monetary score while Europe region has a very small region of high percentage of people connected ot the internet in comparison.


### Tables

```{r}
set.seed(1738)
dt <- (sample(nrow(cleaned_data), nrow(cleaned_data)*.7))
train <- cbind( cleaned_data$Percent_People_Connected[dt], cleaned_data$Monetary_Freedom[dt])
test<- cbind( cleaned_data$Percent_People_Connected[-dt], cleaned_data$Monetary_Freedom[-dt])
true_responses <- as.matrix(cleaned_data[dt, 2])
test_responses <- as.factor(as.matrix(cleaned_data[-dt, 2]))
knnModel <- knn(train = train, test = test, cl=true_responses, k=5)
confMat <- confusionMatrix(data =knnModel, reference=test_responses) 
confMat$table <- rbind(as.matrix(confMat$table), c(10,8,17,4,11))
confMat$table <- cbind(confMat$table, c(5,9,17,1,17, 50))
knitr::kable(confMat$table,col.names = c("Predicted","Americas",	"Asia-Pacific",	"Europe",	"Middle East/North Africa",	"Sub-Saharan Africa","Total"	))
knitr::kable(confMat$byClass[, c(5,11)])

```

Plot {data-navmenu="KNN Classification" data-orientation=columns}
=====================


### Plot

```{r}
plot.df <- data.frame(data=train, Region = true_responses, trained=rep(1,nrow(true_responses)))


plot.df1 = data.frame(x = plot.df$data.1, 
                      y = plot.df$data.2, 
                      Region = plot.df$Region)
find_hull = function(df) df[chull(df$x, df$y), ]
boundary = ddply(plot.df1, .variables = "Region", .fun = find_hull)
plot.df <- rbind(plot.df, data.frame(data=test, Region = test_responses, trained=rep(0,length(test_responses))))
colnames(plot.df) <- c("Percent_People_Connected", "Monetary_Freedom", "Region", "Trained")
ggplot() +
  geom_polygon(data = boundary, aes(x = x, y = y, group = Region, fill = Region), alpha = 0.3) +
  geom_point(data = plot.df, aes(x = Percent_People_Connected, y = Monetary_Freedom, pch = factor(Trained), color=factor(Region)),size=2) +
  labs(x = "Percent People Connected", y = "Monetary Freedom", color = "Region", pch = "Type of Data\n0 for Test\n1for Train",title = "The Clusters Defining Each Region Data") 

```



Naive Bayes Classification {data-orientation=columns}
=====================


### Description
For K-Nearest Neighbors Classification we decided to try and figure out based on Overall Economic Freedom Score and percentage of internet users how you can classify what region that country belongs to. We already have our data classified based on region and some of the regions are larger than others making this knn classification more biased to those countries. We wish to see if the possible relationships that arise in trying to classify the region, like would certain regions would cluster in certain areas like say, higher Monetary Freedom Score index and high internet connection percentage. 

In our testing we decided the other parameter besides percent of people connected to the internet to classify regions with would be the Monetary Freedom Score. For whatever reason this gives our highest accuracy score of 0.5510204 without adding any more variables to be classified with. One possible explanation could be that when compared with percentage of people connected to the internet, the Monetary Freedom Score has the least looking correlation as seen in our multivariate regression tab. 

The plot shows the section that the training set created when it was being trained on with 70% of the whole data set as its training set. Each colored section represents the region's region for classification. So for example Sub-Saharan Africa has a large region that is mainly clustered with a low percentage of people connected to the internet but a fairly high range of monetary score while Europe region has a very small region of high percentage of people connected ot the internet in comparison.


### Results

```{r}
naiveBayesData <- cleaned_data
naiveBayesData$ClassInternet[cleaned_data$Percent_People_Connected>90] <- "A"
naiveBayesData$ClassInternet[cleaned_data$Percent_People_Connected<=90 & cleaned_data$Percent_People_Connected>77] <- "B"
naiveBayesData$ClassInternet[cleaned_data$Percent_People_Connected<=77 & cleaned_data$Percent_People_Connected>50] <- "C"
naiveBayesData$ClassInternet[cleaned_data$Percent_People_Connected<=50] <- "D"

naiveBayesData$ClassInternet <- as.factor(naiveBayesData$ClassInternet)
naiveBayesData$ClassMoney[cleaned_data$Monetary_Freedom<=75 & cleaned_data$Monetary_Freedom>71] <- "B"
naiveBayesData$ClassMoney[cleaned_data$Monetary_Freedom<=71 & cleaned_data$Monetary_Freedom>67] <- "C"
naiveBayesData$ClassMoney[cleaned_data$Monetary_Freedom<=67] <- "D"

naiveBayesData$ClassMoney <- as.factor(naiveBayesData$ClassMoney)

trainIndex <- sort(sample(nrow(naiveBayesData), nrow(naiveBayesData)*.7))
trainData <- naiveBayesData[trainIndex, c("ClassInternet", "ClassMoney")]
testData <- naiveBayesData[-trainIndex, c("ClassInternet", "ClassMoney")]
trainLabels <- naiveBayesData[trainIndex, "Region"]
testLabels <- naiveBayesData[-trainIndex, "Region"]
model <- naiveBayes(Region ~ ClassInternet+ClassMoney,cbind(trainData,trainLabels))

yhat <- predict(model, testData)
library(gmodels)
CrossTable(yhat, as.matrix(testLabels), prop.chisq = FALSE, prop.t = FALSE, prop.r = FALSE, 
           dnn = c("predicted", "actual"))

resultTable <- table(yhat, as.matrix(testLabels))
knitr::kable(resultTable,caption = "Prediction", col.names = c("Actual", "Americas",	"Asia-Pacific",	"Europe",	"Middle East/North Africa",	"Sub-Saharan Africa"))
misclass <- (sum(resultTable) - sum(diag(resultTable))) / sum(resultTable)
#misclass

```


Logistic Regression Classification {data-orientation=rows}
=====================

### Description
Since the data did not contain a data type that is required by Logistic regression, we have to create some new variables. We defined the threshold for Percent People Connected and Tax burden by using mean(), then create a two new variables to record the result. The new created variables are Connected and Heavy. Connected = 1 means the percent of people of this country is greater than the mean value, Connected = 0 means the percent of people of this country is smaller than the mean value. Heavy = Heavy(= 1) means the tax burden of this country is greater than the mean value, Heavy = Not Heavy(= 0) means the tax burden of this country is smaller than the mean value. After we set up the new variables, then we can do the Logistic regression. We used Financial Freedom and Property Rights as predictors and we found that the country have heavy tax burden have more percent people connected to the internet in general. And we also can get two prediction formula to predict, logit(̂ π) = -2.03 + 0.059 Financial Freedom − 1.37 Not Heavy, logit(̂ π) = -1.85 + 0.053 Property Rights − 1.88 Not Heavy


{data-orientation=rows}
------------------

### Results
```{r plot6, height=7, width=5}
library(dplyr)
#since there is no required data type to use Logistic regression, so I need to create some new variables to satisfy it.
#setup the threshold, if the percent people connected is over the mean value than set it as 1, otherwise set it as 0
threshold <- mean(cleaned_data$Percent_People_Connected)
# calculate the mean of tax burden and set it as the threshold of tax burden
mean <- mean(cleaned_data$Tax_Burden)


# Create a binary response variable based on the threshold
cleaned_data <- cleaned_data %>%
  mutate(Connected = ifelse(Percent_People_Connected > threshold, 1, 0))
cleaned_data <- cleaned_data %>%
  mutate(Heavy = ifelse(Tax_Burden > mean, "Heavy", "Not heavy"))

# Do the logistic regression, take Financial freedom and new created variable Heavy as predictors, target variable is new created variable Connected
logit_model <- glm(Connected ~ Financial_Freedom* Heavy, data = cleaned_data, family = binomial)

# plot for logistic regression
ggplot(cleaned_data, aes(x = Financial_Freedom, y = Connected, color = Heavy)) +
  geom_point(aes(color = Heavy), position = position_jitter(height = 0.03, width = 0)) +
  geom_smooth(method = "glm", method.args = list(family="binomial"), se = F) +
  labs(x = "`Financial Freedom`", y = "Connected", title = "Logistic Regression: Connected vs. Property Rights")
#confint.default(logit_model)


# Do the logistic regression, take Property Rights and new created variable Heavy as predictors, target variable is new created variable Connected
logit_model <- glm(Connected ~ Property_Rights* Heavy, data = cleaned_data, family = binomial)
# plot for logistic regression
ggplot(cleaned_data, aes(x = Property_Rights, y = Connected, color = Heavy)) +
  geom_point(aes(color = Heavy), position = position_jitter(height = 0.03, width = 0)) +
  geom_smooth(method = "glm", method.args = list(family="binomial"), se = F) +
  labs(x = "`Property Rights`", y = "Connected", title = "Logistic Regression: Connected vs. Property Rights")
#confint.default(logit_model)

```